---
title: "ckme project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import dataset

```{r}
data <- read.csv("/Users/zafrinasomani/Desktop/CKME 136/mental-health-in-tech-2016.csv", header = T, na.strings = c("", "NA"))
clean <- data
```

Inspect dataset 

```{r}
head(clean)
#tail(clean)
summary(clean)
str(clean)
dim(clean)
head(names(clean))
```

Assign correct data types

```{r}
clean$Are.you.self.employed. <- as.factor(clean$Are.you.self.employed.)
clean$Is.your.employer.primarily.a.tech.company.organization. <- as.factor(clean$Is.your.employer.primarily.a.tech.company.organization.)
clean$Is.your.primary.role.within.your.company.related.to.tech.IT. <- as.factor(clean$Is.your.primary.role.within.your.company.related.to.tech.IT.)
clean$Do.you.have.medical.coverage..private.insurance.or.state.provided..which.includes.treatment.of..mental.health.issues. <- as.factor(clean$Do.you.have.medical.coverage..private.insurance.or.state.provided..which.includes.treatment.of..mental.health.issues.)
clean$Do.you.have.previous.employers. <- as.factor(clean$Do.you.have.previous.employers.)
clean$Have.you.ever.sought.treatment.for.a.mental.health.issue.from.a.mental.health.professional. <- as.factor(clean$Have.you.ever.sought.treatment.for.a.mental.health.issue.from.a.mental.health.professional.)
```

Removed individuals who are self-employed and who do not work in an IT company or in the tech department

```{r}
clean <- subset(clean, clean$Are.you.self.employed. == 0)
clean <- subset(clean, clean$Is.your.primary.role.within.your.company.related.to.tech.IT. == 1 | clean$Is.your.employer.primarily.a.tech.company.organization. == 1)
```

Remove questions questions that were asked only to self-employed individuals. 

```{r}
clean$Do.you.have.medical.coverage..private.insurance.or.state.provided..which.includes.treatment.of..mental.health.issues. <- NULL
clean$Do.you.know.local.or.online.resources.to.seek.help.for.a.mental.health.disorder. <- NULL
clean$If.you.have.been.diagnosed.or.treated.for.a.mental.health.disorder..do.you.ever.reveal.this.to.clients.or.business.contacts. <- NULL
clean$If.you.have.revealed.a.mental.health.issue.to.a.client.or.business.contact..do.you.believe.this.has.impacted.you.negatively. <- NULL
clean$If.you.have.been.diagnosed.or.treated.for.a.mental.health.disorder..do.you.ever.reveal.this.to.coworkers.or.employees. <- NULL
clean$If.you.have.revealed.a.mental.health.issue.to.a.coworker.or.employee..do.you.believe.this.has.impacted.you.negatively. <- NULL
clean$Do.you.believe.your.productivity.is.ever.affected.by.a.mental.health.issue. <- NULL
clean$If.yes..what.percentage.of.your.work.time..time.performing.primary.or.secondary.job.functions..is.affected.by.a.mental.health.issue. <- NULL
```

Remove outliers in age which appear to be data entry errors:

```{r}
clean <- subset(clean, !clean$What.is.your.age. == 323)
clean <- subset(clean, !clean$What.is.your.age. == 3)
clean <- subset(clean, !clean$What.is.your.age. == 99)
```

Distribution of Age 

```{r, echo=FALSE}
boxplot(clean$What.is.your.age., main = "Age Boxplot")
hist(clean$What.is.your.age., main = "Age Histogram", xlab = "Age")
```

Rename columns - shorten column names 

```{r}
names(clean)[names(clean) == "Are.you.self.employed."] <- c("self_employed")
names(clean)[names(clean) == "How.many.employees.does.your.company.or.organization.have."] <- c("no_of_employees")
names(clean)[names(clean) == "Is.your.employer.primarily.a.tech.company.organization."] <- c("tech_company")
names(clean)[names(clean) == "Is.your.primary.role.within.your.company.related.to.tech.IT."] <- c("tech_role")
names(clean)[names(clean) == "Does.your.employer.provide.mental.health.benefits.as.part.of.healthcare.coverage."] <- c("mental_health_benefits")
names(clean)[names(clean) == "Do.you.know.the.options.for.mental.health.care.available.under.your.employer.provided.coverage."] <- c("options_mental_health")
names(clean)[names(clean) == "Has.your.employer.ever.formally.discussed.mental.health..for.example..as.part.of.a.wellness.campaign.or.other.official.communication.."] <- c("wellness_campaign")
names(clean)[names(clean) == "Does.your.employer.offer.resources.to.learn.more.about.mental.health.concerns.and.options.for.seeking.help."] <- c("mh_resources")
names(clean)[names(clean) == "Is.your.anonymity.protected.if.you.choose.to.take.advantage.of.mental.health.or.substance.abuse.treatment.resources.provided.by.your.employer."] <- c("anonymity")
names(clean)[names(clean) == "If.a.mental.health.issue.prompted.you.to.request.a.medical.leave.from.work..asking.for.that.leave.would.be."] <- c("mh_leave_request")
names(clean)[names(clean) == "Do.you.think.that.discussing.a.mental.health.disorder.with.your.employer.would.have.negative.consequences."] <- c("mh_discuss_neg_cons")
names(clean)[names(clean) == "Do.you.think.that.discussing.a.physical.health.issue.with.your.employer.would.have.negative.consequences."] <- c("ph_discuss_neg_cons")
names(clean)[names(clean) == "Would.you.feel.comfortable.discussing.a.mental.health.disorder.with.your.coworkers."] <- c("mh_discuss_coworker")
names(clean)[names(clean) == "Would.you.feel.comfortable.discussing.a.mental.health.disorder.with.your.direct.supervisor.s.."] <- c("mh_discuss_supervisor")
names(clean)[names(clean) == "Do.you.feel.that.your.employer.takes.mental.health.as.seriously.as.physical.health."] <- c("mh_serious_ph")
names(clean)[names(clean) == "Have.you.heard.of.or.observed.negative.consequences.for.co.workers.who.have.been.open.about.mental.health.issues.in.your.workplace."] <- c("observe_negcons_mh")
names(clean)[names(clean) == "Do.you.have.previous.employers."] <- c("previous_employer")
names(clean)[names(clean) == "Have.your.previous.employers.provided.mental.health.benefits."] <- c("pe_mh_benefits")
names(clean)[names(clean) == "Were.you.aware.of.the.options.for.mental.health.care.provided.by.your.previous.employers."] <- c("pe_mh_options")
names(clean)[names(clean) == "Did.your.previous.employers.ever.formally.discuss.mental.health..as.part.of.a.wellness.campaign.or.other.official.communication.."] <- c("pe_wellness_campaign")
names(clean)[names(clean) == "Did.your.previous.employers.provide.resources.to.learn.more.about.mental.health.issues.and.how.to.seek.help."] <- c("pe_mh_resources")
names(clean)[names(clean) == "Was.your.anonymity.protected.if.you.chose.to.take.advantage.of.mental.health.or.substance.abuse.treatment.resources.with.previous.employers."] <- c("pe_anonymity")
names(clean)[names(clean) == "Do.you.think.that.discussing.a.mental.health.disorder.with.previous.employers.would.have.negative.consequences."] <- c("pe_mh_discuss_negcons")
names(clean)[names(clean) == "Do.you.think.that.discussing.a.physical.health.issue.with.previous.employers.would.have.negative.consequences."] <- c("pe_ph_discuss_negcons")
names(clean)[names(clean) == "Would.you.have.been.willing.to.discuss.a.mental.health.issue.with.your.previous.co.workers."] <- c("pe_mh_discuss_coworker")
names(clean)[names(clean) == "Would.you.have.been.willing.to.discuss.a.mental.health.issue.with.your.direct.supervisor.s.."] <- c("pe_mh_discuss_supervisor")
names(clean)[names(clean) == "Did.you.feel.that.your.previous.employers.took.mental.health.as.seriously.as.physical.health."] <- c("pe_mh_serious_ph")
names(clean)[names(clean) == "Did.you.hear.of.or.observe.negative.consequences.for.co.workers.with.mental.health.issues.in.your.previous.workplaces."] <- c("pe_observe_negcons_mh")
names(clean)[names(clean) == "Would.you.be.willing.to.bring.up.a.physical.health.issue.with.a.potential.employer.in.an.interview."] <- c("ph_interview")
names(clean)[names(clean) == "Why.or.why.not."] <- c("ph_interview_why")
names(clean)[names(clean) == "Would.you.bring.up.a.mental.health.issue.with.a.potential.employer.in.an.interview."] <- c("mh_interview")
names(clean)[names(clean) == "Why.or.why.not..1"] <- c("mh_interview_why")
names(clean)[names(clean) == "Do.you.feel.that.being.identified.as.a.person.with.a.mental.health.issue.would.hurt.your.career."] <- c("id_mh_hurt_career")
names(clean)[names(clean) == "Do.you.think.that.team.members.co.workers.would.view.you.more.negatively.if.they.knew.you.suffered.from.a.mental.health.issue."] <- c("team_neg_view_mh")
names(clean)[names(clean) == "How.willing.would.you.be.to.share.with.friends.and.family.that.you.have.a.mental.illness."] <- c("friends_family")
names(clean)[names(clean) == "Have.you.observed.or.experienced.an.unsupportive.or.badly.handled.response.to.a.mental.health.issue.in.your.current.or.previous.workplace."] <- c("unsupportive_response_mh")
names(clean)[names(clean) == "Have.your.observations.of.how.another.individual.who.discussed.a.mental.health.disorder.made.you.less.likely.to.reveal.a.mental.health.issue.yourself.in.your.current.workplace."] <- c("obs_impact")
names(clean)[names(clean) == "Do.you.have.a.family.history.of.mental.illness."] <- c("mh_family_history")
names(clean)[names(clean) == "Have.you.had.a.mental.health.disorder.in.the.past."] <- c("mh_past")
names(clean)[names(clean) == "Do.you.currently.have.a.mental.health.disorder."] <- c("mh_current")
names(clean)[names(clean) == "If.yes..what.condition.s..have.you.been.diagnosed.with."] <- c("current_diagnosis")
names(clean)[names(clean) == "If.maybe..what.condition.s..do.you.believe.you.have."] <- c("condition_undiagnosed")
names(clean)[names(clean) == "Have.you.been.diagnosed.with.a.mental.health.condition.by.a.medical.professional."] <- c("diagnosed_medpro")
names(clean)[names(clean) == "If.so..what.condition.s..were.you.diagnosed.with."] <- c("condition_diagnosed_medpro")
names(clean)[names(clean) == "Have.you.ever.sought.treatment.for.a.mental.health.issue.from.a.mental.health.professional."] <- c("sought_treatment")
names(clean)[names(clean) == "If.you.have.a.mental.health.issue..do.you.feel.that.it.interferes.with.your.work.when.being.treated.effectively."] <- c("mhi_work_treated_effectively")
names(clean)[names(clean) == "If.you.have.a.mental.health.issue..do.you.feel.that.it.interferes.with.your.work.when.NOT.being.treated.effectively."] <- c("mhi_work_not_treated_effectively")
names(clean)[names(clean) == "What.is.your.age."] <- c("age")
names(clean)[names(clean) == "What.is.your.gender."] <- c("gender")
names(clean)[names(clean) == "What.country.do.you.live.in."] <- c("home_country")
names(clean)[names(clean) == "What.US.state.or.territory.do.you.live.in."] <- c("home_state")
names(clean)[names(clean) == "What.country.do.you.work.in."] <- c("work_country")
names(clean)[names(clean) == "What.US.state.or.territory.do.you.work.in."] <- c("work_state")
names(clean)[names(clean) == "Which.of.the.following.best.describes.your.work.position."] <- c("work_position")
names(clean)[names(clean) == "Do.you.work.remotely."] <- c("work_remotely")
summary(clean)
```

Clean gender variable

```{r}
clean$gender <- as.character(clean$gender)
clean$gender[clean$gender == "Agender"] <- c("other")
clean$gender[clean$gender == "Androgynous"] <- c("other")
clean$gender[clean$gender == "Bigender"] <- c("other")
clean$gender[clean$gender == "Cis female"] <- c("female")
clean$gender[clean$gender == "cis male"] <- c("male")
clean$gender[clean$gender == "Cis male"] <- c("male")
clean$gender[clean$gender == "Cis Male"] <- c("male")
clean$gender[clean$gender == "cis man"] <- c("male")
clean$gender[clean$gender == "Cis-woman"] <- c("female")
clean$gender[clean$gender == "cisdude"] <- c("male")
clean$gender[clean$gender == "Cisgender Female"] <- c("female")
clean$gender[clean$gender == "Dude"] <- c("male")
clean$gender[clean$gender == "Enby"] <- c("other")
clean$gender[clean$gender == "f"] <- c("female")
clean$gender[clean$gender == "F"] <- c("female")
clean$gender[clean$gender == "fem"] <- c("female")
clean$gender[clean$gender == "Female"] <- c("female")
clean$gender[clean$gender == "female "] <- c("female")
clean$gender[clean$gender == "Female "] <- c("female")
clean$gender[clean$gender == "Female (props for making this a freeform field, though)"] <- c("female")
clean$gender[clean$gender == "Female assigned at birth"] <- c("female")
clean$gender[clean$gender == "Female or Multi-Gender Femme"] <- c("female")
clean$gender[clean$gender == "female-bodied; no feelings about gender"] <- c("female")
clean$gender[clean$gender == "female/woman"] <- c("female")
clean$gender[clean$gender == "fm"] <- c("female")
clean$gender[clean$gender == "Genderfluid "] <- c("other")
clean$gender[clean$gender == "Genderfluid (born female)"] <- c("other")
clean$gender[clean$gender == "Genderflux demi-girl"] <- c("other")
clean$gender[clean$gender == "genderqueer"] <- c("other")
clean$gender[clean$gender == "Genderqueer"] <- c("other")
clean$gender[clean$gender == "genderqueer woman"] <- c("other")
clean$gender[clean$gender == "Human"] <- c("other")
clean$gender[clean$gender == "I identify as female."] <- c("female")
clean$gender[clean$gender == "m"] <- c("male")
clean$gender[clean$gender == "M"] <- c("male")
clean$gender[clean$gender == "mail"] <- c("male")
clean$gender[clean$gender == "male "] <- c("male")
clean$gender[clean$gender == "Male"] <- c("male")
clean$gender[clean$gender == "Male "] <- c("male")
clean$gender[clean$gender == "Male (cis)"] <- c("male")
clean$gender[clean$gender == "Male (trans, FtM)"] <- c("male")
clean$gender[clean$gender == "male 9:1 female, roughly"] <- c("other")
clean$gender[clean$gender == "Male."] <- c("male")
clean$gender[clean$gender == "Male/genderqueer"] <- c("other")
clean$gender[clean$gender == "man"] <- c("male")
clean$gender[clean$gender == "Man"] <- c("male")
clean$gender[clean$gender == "N/A"] <- c("other")
clean$gender[clean$gender == "nb masculine"] <- c("other")
clean$gender[clean$gender == "non-binary"] <- c("other")
clean$gender[clean$gender == "Nonbinary"] <- c("other")
clean$gender[clean$gender == "none of your business"] <- c("other")
clean$gender[clean$gender == "Other/Transfeminine"] <- c("other")
clean$gender[clean$gender == "Queer"] <- c("other")
clean$gender[clean$gender == "Sex is male"] <- c("male")
clean$gender[clean$gender == "Transgender woman"] <- c("other")
clean$gender[clean$gender == "Transitioned, M2F"] <- c("female")
clean$gender[clean$gender == "Woman"] <- c("female")
clean$gender[clean$gender == "Cis female "] <- c("female")
clean$gender[clean$gender == "Female assigned at birth "] <- c("female")
clean$gender[clean$gender == "Genderfluid"] <- c("other")
clean$gender[clean$gender == "woman"] <- c("female")
clean$gender[is.na(clean$gender)] <- c("other")
clean$gender <- as.factor(clean$gender)
summary(clean$gender)
```

Create new dataframe with current diagnoses - will show individuals with multiple diagnoses and what the diagnoses are 

```{r}
library(tidyr)
library(dplyr)
currentdiagnosis <- clean %>% separate(current_diagnosis, sep = '\\|', c("curdiag_1", "curdiag_2", "curdiag_3","curdiag_4","curdiag_5","curdiag_6","curdiag_7","curdiag_8"), fill = 'right')
currentdiagnosis %>% select(matches('curdiag_[1-8]')) %>% str()
currentdiagnosis <- currentdiagnosis %>% select(matches("curdiag_[1-8]")) %>% mutate_each(funs = 'as.factor')
currentdiagnosis %>% select(matches("curdiag_[1-8]")) %>% str()
currentdiagnosis$na_count <- as.factor(apply(is.na(currentdiagnosis), MARGIN = 1, FUN = sum))
summary(currentdiagnosis)
```

Create new dataframe separating work positions, so each work position is in a new column

```{r}
clean$work_position <- as.factor(clean$work_position)
job <- clean %>% separate(work_position, sep = '\\|', c("wp_1", "wp_2", "wp_3", "wp_4", "wp_5", "wp_6", "wp_7", "wp_8", "wp_9", "wp_10"), fill = 'right')
job %>% select(matches('wp_[1-9]', 'wp_10')) %>% str()
job <- job %>% select(matches("wp_[1-9]", "wp_10")) %>% mutate_each(funs = 'as.factor')
job %>% select(matches("wp_[1-9]", "wp_10")) %>% str()
job$na_count <- as.factor(apply(is.na(job), MARGIN = 1, FUN = sum))
summary(job)
```

Create new dataframe separating unconditioned diagnoses

```{r}
undiagnosed <- clean %>% separate(condition_undiagnosed, sep = '\\|', c("uc_1", "uc_2", "uc_3", "uc_4", "uc_5", "uc_6", "uc_7", "uc_8"), fill = 'right')
undiagnosed %>% select(matches('uc_[1-8]')) %>% str()
undiagnosed <- undiagnosed %>% select(matches('uc_[1-8]')) %>% mutate_each(funs = 'as.factor')
undiagnosed %>% select(matches("uc_[1-8]")) %>% str()
undiagnosed$na_count <- as.factor(apply(is.na(undiagnosed), MARGIN = 1, FUN = sum))
summary(undiagnosed)
```

Create new dataframe separating diagnosed medication conditions

```{r}
med_diagnosis <- clean %>% separate(condition_diagnosed_medpro, sep = '\\|', c("cond_1", "cond_2", "cond_3", "cond_4", "cond_5", "cond_6"), fill = 'right')
med_diagnosis %>% select(matches('cond_[1-6]')) %>% str()
med_diagnosis <- med_diagnosis %>% select(matches('cond_[1-6]')) %>% mutate_each(funs = 'as.factor')
med_diagnosis %>% select(matches('cond_[1-6]')) %>% str()
med_diagnosis$na_count <- as.factor(apply(is.na(med_diagnosis), MARGIN = 1, FUN = sum))
summary(med_diagnosis)
```

Clean med_diagnosis levels: 

```{r}
med_diagnosis$cond_1 <- as.character(med_diagnosis$cond_1)
med_diagnosis$cond_1[med_diagnosis$cond_1 == "attention deficit disorder (but not the hyperactive version)"] <- c("ADD/ADHD")
med_diagnosis$cond_1[med_diagnosis$cond_1 == "Attention Deficit Hyperactivity Disorder"] <- c("ADD/ADHD")
med_diagnosis$cond_1[med_diagnosis$cond_1 == "ADD (w/o Hyperactivity)"] <- c("ADD/ADHD")
med_diagnosis$cond_1[med_diagnosis$cond_1 == "Autism - while not a \"mental illness\", still greatly affects how I handle anxiety"] <- c("Autism Spectrum Disorder")
med_diagnosis$cond_1[med_diagnosis$cond_1 == "Autism (Asperger's)"] <- c("Asperger Syndrome")
med_diagnosis$cond_1[med_diagnosis$cond_1 == "autism spectrum disorder"] <- c("Autism Spectrum Disorder")
med_diagnosis$cond_1 <- as.factor(med_diagnosis$cond_1)
```


See how many people live and work in the same country 

```{r}
clean$country <- as.factor(ifelse(as.character(clean$home_country) == as.character(clean$work_country), c("same"), c("different")))
summary(clean$country)
```

See how many people live and work in the same state (applicable to individual working in the US only) - NA refers to individuals who do not live in the US

```{r}
clean$state <- as.factor(ifelse(as.character(clean$home_state) == as.character(clean$work_state), c("same"), c("different")))
summary(clean$state)
```

EXPLORATORY ANALYSIS 

Most common current diagnoses among IT/tech workers:

```{r}
library(ggplot2)
subcd1 <- subset(currentdiagnosis, !is.na(currentdiagnosis$curdiag_1))
ggplot(data = subcd1, aes(x = curdiag_1)) + geom_bar(fill = 'red') + labs(x = "Diagnosis", y = "Count") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Most common diagnosis among IT/Tech workers are Anxety and Mood disorders. On average, most individuals disclosed one diagnosis, however some individuals stated up to a combination of 8 diagnoses. 

Most common work position: 

```{r}
wp1 <- subset(job, !is.na(job$wp_1))
ggplot(data = wp1, aes(x = wp_1)) + geom_bar(fill = 'purple') + labs(x = "Position", y = "Count") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The most common work positions are Back-end Developer and Front-end Developer. Most individuals stated they did at least one job, with individuals stating they did up to 10 different jobs. 

Most common undiagnosed condition : 

```{r}
undiag1 <- subset(undiagnosed, !is.na(undiagnosed$uc_1))
ggplot(data = undiag1, aes(x = uc_1)) + geom_bar(fill = 'green') + labs(x = "Undiagnosed Condition", y = "Count") + theme(axis.text.x = element_text(angle = 90))
```

Most individuals feel they have an undiagnosed Anxiety or Mood disorder. 

Most common medically diagnosed condition: 

```{r}
meddiag1 <- subset(med_diagnosis, !is.na(med_diagnosis$cond_1))
ggplot(data = meddiag1, aes(x = cond_1)) + geom_bar(fill = 'yellow') + labs(x = "Medical Diagnosis", y = "Count") + theme(axis.text.x = element_text(angle = 90))
```

Most individuals have been medically diagnosed with an Anxiety or Mood disorder 

Demographic Data: 

```{r}
ggplot(data = clean, aes(x = home_country)) + geom_bar(fill = 'blue') + labs(x = "Country", y = "Count") + theme(axis.text.x = element_text(angle = 90))
ggplot(data = clean, aes(x = work_country)) + geom_bar(fill = 'blue') + labs(x = "Country", y = "Count") + theme(axis.text.x = element_text(angle = 90))
```

Most respondents are from the United States and work in the United States, with the UK coming in second. 

Most respondents live and work in the same country. 

```{r}
ggplot(data = clean, aes(x = country)) + geom_bar(fill = 'blue') + labs(x = "Country: Work and Live", y = "Count")
```

Gender of Repondents

```{r}
ggplot(data = clean, aes(x = gender)) + geom_bar(fill = 'orange') + labs(x = "Gender", y = "Count")
```

Create new dataframe, removing attributes that have been moved to different dataframes 

```{r}
analysis <- clean
analysis$ph_interview_why <- NULL 
analysis$mh_interview_why <- NULL 
analysis$current_diagnosis <- NULL
analysis$condition_undiagnosed <- NULL 
analysis$condition_diagnosed_medpro <- NULL
analysis$work_position <- NULL 
analysis$home_country <- NULL
analysis$home_state <- NULL 
analysis$work_country <- NULL
analysis$work_state <- NULL
analysis$self_employed <- NULL
```

Clean up 'analysis' 

```{r}
analysis$tech_role <- factor(clean$tech_role, levels = c("O", "1"), labels = c("No", "Yes"))
analysis$sought_treatment <- factor(analysis$sought_treatment, levels = c("0", "1"), labels = c("No", "Yes"))
analysis$agegroup <- cut(clean$age, c(9, 19, 29, 39, 49, 59, 69, 79), labels = c("10-19","20-29", "30-39", "40-49", "50-59", "60-69", "70-79"))
summary(analysis$agegroup)
analysis$age <- NULL
```

Dimensionality Reduction 

```{r}
library(FSelector)
as.simple.formula(cutoff.biggest.diff(chi.squared(formula = diagnosed_medpro ~ ., data = analysis)), "diagnosed_medpro")
as.simple.formula(cutoff.biggest.diff(chi.squared(formula = sought_treatment ~ ., data = analysis)), "sought_treatment")
as.simple.formula(cutoff.biggest.diff(information.gain(formula = diagnosed_medpro ~ ., data = analysis)), "diagnosed_medpro")
as.simple.formula(cutoff.biggest.diff(information.gain(formula = sought_treatment ~ ., data = analysis)), "sought_treatment")
as.simple.formula(cfs(formula = diagnosed_medpro ~ ., data = analysis), "diagnosed_medpro")
as.simple.formula(cfs(formula = sought_treatment ~ ., data = analysis), "sought_treatment")
as.simple.formula(consistency(formula = diagnosed_medpro ~ ., data = analysis), "diagnosed_medpro")
as.simple.formula(consistency(formula = sought_treatment ~ ., data = analysis), "sought_treatment")
```

The attributes selected by chi-squared were the same as those that were found using information gain. CFS had attributes that were found in chi-squared. Therefore, attributes found with chi-squared will be used to for model as these attributes were found using three different methods (shows strength). 

Split data into train and test data 

```{r}
set.seed(789)
analysis_train <- sample(nrow(analysis), floor(nrow(analysis)*0.7))
at_train <- analysis[analysis_train,]
at_test <- analysis[-analysis_train,]

class_error_calc <- function(actual, predicted) {
  mean(actual != predicted)
}
```

Logistic Regression Model for Diagnosed by a Medical Professional 

```{r}
dm_lm <- glm(diagnosed_medpro ~ mh_past + sought_treatment + mhi_work_treated_effectively + mhi_work_not_treated_effectively + mh_current, family = "binomial", data = at_train)

summary(dm_lm)
coef(dm_lm)
#betas for the equations:
head(predict(dm_lm))
#probabilities: 
head(predict(dm_lm, type = "response"))
#classifications:
dm_lm_predict <- ifelse(predict(dm_lm, newdata = at_test, type = "link") > 0, "Yes", "No")
table(dm_lm_predict, at_test$diagnosed_medpro)
class_error_calc(actual = at_test$diagnosed_medpro, predicted = dm_lm_predict)
```

Logistic Regression Model for whether or not an individual sought treatment for a mental health condition 
```{r}
st_lm <- glm(sought_treatment ~ diagnosed_medpro + mh_past + mhi_work_treated_effectively + mhi_work_not_treated_effectively + mh_current, family = "binomial", data = at_train)

summary(st_lm)
coef(st_lm)
#betas for the equations:
head(predict(st_lm))
#probabilities 
head(predict(st_lm, type = "response"))
#classifications 
st_lm_predict <- ifelse(predict(st_lm, newdata = at_test, type = "link") > 0, "Yes", "No")
table(st_lm_predict, at_test$sought_treatment)
class_error_calc(actual = at_test$sought_treatment, predicted = st_lm_predict)
```

Create matrix for ANN diganosed by medical professional model with chi-square attributes

```{r}
library(neuralnet)
mm_dm <- model.matrix(~diagnosed_medpro + mh_past + sought_treatment + mhi_work_treated_effectively + mhi_work_not_treated_effectively + mh_current, data = analysis)
head(mm_dm)
```

Create random sample and split matrix data into train and test data 
 
```{r}
set.seed(789)
index_analysis_dm <- sample(nrow(mm_dm), floor(nrow(mm_dm)*0.7))
at_train_dm <- mm_dm[index_analysis_dm,]
at_test_dm <- as.data.frame(mm_dm[-index_analysis_dm,])
```

Artifical Neural Network (ANN) Model for whether or not an individual has been diagnosed with a mental health condition by a medical professional 

```{r}
nn_dm <- neuralnet(diagnosed_medproYes ~ mh_pastNo + mh_pastYes + sought_treatmentYes + mhi_work_treated_effectivelyOften + mhi_work_treated_effectivelyRarely + mhi_work_treated_effectivelySometimes + mhi_work_not_treated_effectivelyOften + mhi_work_not_treated_effectivelyRarely + mhi_work_not_treated_effectivelySometimes + mh_currentNo + mh_currentYes, data = at_train_dm)
plot(nn_dm)

at_test_dm_subset <- subset(at_test_dm, select = c("mh_pastNo", "mh_pastYes", "sought_treatmentYes", "mhi_work_treated_effectivelyOften", "mhi_work_treated_effectivelyRarely", "mhi_work_treated_effectivelySometimes", "mhi_work_not_treated_effectivelyOften", "mhi_work_not_treated_effectivelyRarely", "mhi_work_not_treated_effectivelySometimes", "mh_currentNo", "mh_currentYes"))
nn_dm_predict <- compute(nn_dm, at_test_dm_subset)
nn_dm_results <- data.frame(actual = at_test_dm$diagnosed_medproYes, prediction = nn_dm_predict$net.result)
head(nn_dm_results)
```

Create matrix for ANN sought treatment model with chi-square attributes

```{r}
mm_st <- model.matrix(~sought_treatment + diagnosed_medpro + mh_past + mhi_work_treated_effectively + mhi_work_not_treated_effectively + mh_current, data = analysis)
head(mm_st)
```

Create random sample and split matrix data into train and test

```{r}
index_analysis_st <- sample(nrow(mm_st), floor(nrow(mm_st)*0.7))
at_train_st <- mm_st[index_analysis_st,]
at_test_st <- as.data.frame(mm_st[-index_analysis_st,])
```

Artifical Neural Network (ANN) Model for whether or not an individual sought treatment for a mental health condition 

```{r}
nn_st <- neuralnet(sought_treatmentYes ~ diagnosed_medproYes + mh_pastNo + mh_pastYes + mhi_work_treated_effectivelyOften + mhi_work_treated_effectivelyRarely + mhi_work_treated_effectivelySometimes + mhi_work_not_treated_effectivelyOften + mhi_work_not_treated_effectivelyRarely + mhi_work_not_treated_effectivelySometimes + mh_currentNo + mh_currentYes, data = at_train_st)
plot(nn_st)

at_test_st_subset <- subset(at_test_st, select = c("diagnosed_medproYes", "mh_pastNo", "mh_pastYes", "mhi_work_treated_effectivelyOften", "mhi_work_treated_effectivelyRarely", "mhi_work_treated_effectivelySometimes", "mhi_work_not_treated_effectivelyOften", "mhi_work_not_treated_effectivelyRarely", "mhi_work_not_treated_effectivelySometimes", "mh_currentNo", "mh_currentYes"))
nn_st_predict <- compute(nn_st, at_test_st_subset)
nn_st_results <- data.frame(actual = at_test_st$sought_treatmentYes, prediction = nn_st_predict$net.result)
head(nn_st_results)
```

Decision Tree for whether or not an individual has been diagnosed by a medical professional 

```{r}
library(partykit)

ctree_dm <- ctree(diagnosed_medpro ~ ., data = at_train)
print(ctree_dm)
plot(ctree_dm, type = "simple")

ctree_dm_predict <- predict(ctree_dm, at_test)
table(ctree_dm_predict, at_test$diagnosed_medpro)
```

Decision Tree for whether or not an individual sought treatment for a mental health condtion 

```{r}
ctree_st <- ctree(sought_treatment ~ ., data = at_train)
print(ctree_st)
plot(ctree_st, type = "simple")

ctree_st_predict <- predict(ctree_st, at_test)
table(ctree_st_predict, at_test$sought_treatment)
```

EVALUATION OF MODELS

Confusion Matrices: 

```{r}
library(caret)

dm_lm_predict <- as.factor(dm_lm_predict)
st_lm_predict <- as.factor(st_lm_predict)

confusionMatrix(data = dm_lm_predict, reference = at_test$diagnosed_medpro, positive = c("Yes"))
confusionMatrix(data = st_lm_predict, reference = at_test$sought_treatment, positive = c("Yes"))

confusionMatrix(data = ctree_dm_predict, reference = at_test$diagnosed_medpro, positive = c("Yes"))
confusionMatrix(data = ctree_st_predict, reference = at_test$sought_treatment, positive = c("Yes"))

#confusion matrix for ANN diagnosed medpro
nn_dm_results_round<- as.data.frame(sapply(nn_dm_results,round,digits=0))
table(nn_dm_results_round$actual, nn_dm_results_round$prediction)

#confusion matrix for ANN sought treatment 
nn_st_results_round<- as.data.frame(sapply(nn_st_results,round,digits=0))
table(nn_st_results_round$actual, nn_st_results_round$prediction)
```

Brier Score 

```{r}
library(DescTools)
BrierScore(dm_lm)
BrierScore(st_lm)


ctree_dm_predict <- as.numeric(factor(ctree_dm_predict, levels = c("No", "Yes"), labels = c(0 ,1)))
at_test$diagnosed_medpro <- as.integer(at_test$diagnosed_medpro)
BrierScore(ctree_dm_predict, at_test$diagnosed_medpro)

ctree_st_predict <- as.numeric(factor(ctree_st_predict, levels = c("No", "Yes"), labels = c(0,1)))
at_test$sought_treatment <- as.numeric(at_test$sought_treatment)
BrierScore(ctree_st_predict, at_test$sought_treatment)

BrierScore(nn_dm_results_round$prediction, nn_dm_results_round$actual)
BrierScore(nn_st_results_round$prediction, nn_st_results_round$actual)

```
